---
title: "Final_Mark_Down"
author: "Quinn Bankson"
date: "2023-05-04"
output:
  pdf_document: default
  html_document: default
---

# DarkOrchid4 Final Mark Down 
## How does proximity to coal fired power plant affect health outcomes in the United States?

```{r setup}
```

```{r}
library(tidyverse)
library(ggplot2)
library(moderndive)
library(GGally)
library(janitor)
library(dplyr)
library(geosphere)
library(readr)
```

## Introducing coal plant data to undergo feature engineering. 

```{r}
#Coal plant data was cleaned in an individual folder, the excel was converted to csv and excess worksheets were removed from the workbook. 

coalplants <- read_csv("https://raw.githubusercontent.com/UNCPublicPolicy/final-team-projects-darkorchid/main/DarkOrchid/DATA/Clean/coal_plants_CLEAN.CSV")
```

```{r}
colnames(coalplants)
coalplants <- clean_names(coalplants)
coalplants <- coalplants %>% filter(country == "United States")
coalplants <- coalplants %>% select(parent_id, unit, subnational_unit_province_state, plant, status, year, local_area_taluk_county, latitude, longitude )
coalplants <- rename(coalplants, county = local_area_taluk_county)
```

```{r}
counties <- read_csv(file = "https://raw.githubusercontent.com/UNCPublicPolicy/final-team-projects-darkorchid/main/DarkOrchid/DATA/Clean/uscounties_CLEAN.csv")
head(counties)
colnames(counties)
county_center <- counties %>% select(county, state_name, lat, lng)
```

```{r}
coalplants_op <- coalplants %>% filter(status == "operating")
coalplants_op <- distinct(coalplants_op, plant, .keep_all = TRUE)
```

```{r}
coal_plants <- coalplants_op %>% select("latitude", "longitude","subnational_unit_province_state", "plant", "county")
coal_plants <- rename(coal_plants, long= longitude, lat = latitude)
coal_plants <- rename(coal_plants, state = subnational_unit_province_state, name = plant)
county_centroids <- county_center
county_centroids <- rename(county_centroids, state = state_name)
```

```{r}
### "latitude", "longitude","subnational_unit_province_state", "plant", "county" were selected as the most important values of interest for the rest of the project. The names were changed but these are the orginal values of interest as they were known in the orignal coal plant data set from globalcoaltracker. 
```

## Feature engineering - calculating the minimum distance from every county centroid in the US to the nearest coal fired power plant using the Haversine formula and longitudinal and latitudinal data. 

```{r}
#install.packages("geosphere")
library(geosphere)
library(dplyr)
library(naniar)

county_centroids <- rename(county_centroids, long = lng)

coal_plants <- coal_plants %>% 
  filter(!is.na(long))
coal_plants <- coal_plants %>% 
  filter(!is.na(lat))

county_centroids <- county_centroids %>% 
  filter(!is.na(long))
county_centroids <- county_centroids %>% 
  filter(!is.na(lat))
```

```{r}
library(dplyr)
library(geosphere)

# filtering to be sure
coal_plants <- na.omit(coal_plants)

distances <- geosphere::distm(county_centroids[, c("long", "lat")], coal_plants[, c("long", "lat")])

# minimum distance per county
min_distances <- apply(distances, 1, min)

# add the new column to the county_centroids dataframe using mutate
county_centroids <- county_centroids %>% 
  mutate(distance_to_nearest_plant = min_distances)

coords_distance <- county_centroids
```

```{r, warning = FALSE}
# filtering to be sure, this code will atatch plant name as well as the minimum distance
coal_plants <- na.omit(coal_plants)

 #repeating earlier to be sure
distances <- geosphere::distm(county_centroids[, c("long", "lat")], coal_plants[, c("long", "lat")])

# minimum distance and corresponding plant name per county
min_distances <- apply(distances, 1, min)
names_of_nearest_plants <- apply(distances, 1, function(x) coal_plants$name[which.min(x)])

# add new columns to the county_centroids dataframe using mutate
county_centroids <- county_centroids %>% 
  mutate(distance_to_nearest_plant = min_distances,
         name_of_nearest_plant = names_of_nearest_plants)

#file_path1 <- "https://raw.githubusercontent.com/UNCPublicPolicy/final-team-projects-darkorchid/main/DarkOrchid/DATA/Clean/centroids_plants_CLEAN.csv"

### A note about write_csv in this .Rmd: all write_csv commands are tagged out to avoid problems with the online repository. The files that are written from this .Rmd were created locally and pushed into their proper place in the repository. If any user of this .Rmd plans to use write_csv to write csvs into their own local pathway, please replace the #file_pathX pathway with your own local path. ###

#file_path1 <- "https://raw.githubusercontent.com/UNCPublicPolicy/final-team-projects-darkorchid/main/DarkOrchid/DATA/Clean/centroids_plants_CLEAN.csv"

#write_csv(county_centroids, file_path1)

```

```{r}
summary(county_centroids$distance_to_nearest_plant)
```

```{r}
### This concludes the feature engineering in which the minimum distance from each county centroid to the nearest coal fired power plant is calculated and compiled along with the name of the nearest plant into a tibble called countycoords.csv. This csv has record of 3143 US counties and their distance. ### 
```

## The next section of the .rmd will compile 3196 health outcomes (3143 of which correspond to the 3143 US counties that are observed) and aqi data from 1036 observed US counties as well. A crosswalk was used to match values by FIPS codes. 

```{r libraries}
library(tidyverse)
library(ggplot2)
library(moderndive)
library(GGally)
library(janitor)
library(dplyr)
library(stringr)
#install.packages("tidyr")
library(tidyr)
```

```{r read files}
crossw <- read_csv(file = "https://raw.githubusercontent.com/UNCPublicPolicy/final-team-projects-darkorchid/main/DarkOrchid/DATA/Clean/countycrosswalk_CLEAN.csv")

aqi <- read_csv(file = "https://raw.githubusercontent.com/UNCPublicPolicy/final-team-projects-darkorchid/main/DarkOrchid/DATA/Clean/aqi_2014_CLEAN.csv")

outcomes <- read_csv(file = "https://raw.githubusercontent.com/UNCPublicPolicy/final-team-projects-darkorchid/main/DarkOrchid/DATA/Clean/health_outcomes_CLEAN.csv")
```


```{r correct names}

colnames(crossw)[1] <- "County"
colnames(crossw)[2] <- "State"
colnames(crossw)[3] <- "FIPS"

crossw <- crossw %>% select(County, State, FIPS)

crossw <- crossw %>%
  filter(County != "County Name")
```

```{r}

crossw <- crossw %>%
  mutate(County = str_to_title(tolower(County)))
```


```{#r}
outcomes <- outcomes %>%
  separate(Location, into = c("County", "State"), sep = ", ", remove = FALSE)

outcomes <- outcomes %>%
  filter(!is.na(State))

outcomes$County <- str_replace(outcomes$County, " County", "")

outcomes$County <- str_replace(outcomes$County, " Parish", "")

outcomes$County <- str_replace(outcomes$County, "Saint ", "St. ")
```

```{r}
# join AQI and Outcomes, there will be all 3000+ Outcomes visible and only 1000ish AQI present

joined_dv_iv <- left_join(outcomes, aqi, by = c("County" = "County", "State" = "State"))
view(joined_dv_iv)
```


```{r}
joined_dv_iv <- clean_names(joined_dv_iv)

# I intend to join the countycoords tibble which has county, state, lat, long, distance_to_nearest_plant, name_of_nearest_plant 

centroidsandplants <- county_centroids

```

```{r}
final <- left_join(joined_dv_iv, centroidsandplants, by = c("county" = "county", "state" = "state"))
view(final)
```



```{r}
library(naniar)

vis_miss(final)
```

```{r}

#file_path2 <- UNCPublicPolicy/final-team-projects-darkorchid/DATA/clean/health_aqi_coal_distance_CLEAN.csv

#write_csv(final, file_path2)

```


```{r}
colnames(final)

```

```{r}
### At this point in the process, we had a merged data set containing health outcomes, air quality, and distance from operating coal plants by county for 3143 counties in the United States. We established health outcomes as our dependent variables ("mortality_resp", "perc_mortality_change_resp", "mortality_ast", "perc_mortality_change_ast", "mortality_lung", "perc_mortality_change_lung"). The remaining variables were independent variables and controls of interest to us. We began conducting some basic regressions. ###
```

## Basic Regressions

```{r}
library(tidyverse)
library(ggplot2)
library(moderndive)
library(GGally)
library(janitor)
library(dplyr)
library(stringr)
library(tidyr)
library(naniar)
```

```{r}
data <- final
```

```{r}
head(data)
colnames(data)
```

```{r}
# Checking to see which variables would create a regression using 1000+ observations and which variables would create a regression using 3000+ observations. 

vis_miss(data)
```

```{r}
data <- data %>% 
  mutate(distance_km = distance_to_nearest_plant / 1000)

pr1model <- lm(mortality_resp_2014 ~ median_aqi + unhealthy_days + hazardous_days + days_no2 + distance_km, data = data)
summary(pr1model)
```

```{r}
pr2model <- lm(mortality_resp_2014 ~ median_aqi + max_aqi + good_days + moderate_days + unhealthy_for_sensitive_groups_days + unhealthy_days + hazardous_days +days_ozone + days_no2 + distance_km, data = data)
summary(pr2model)
```

```{r}

### After observing some basic regressions, we found our R-squared values to be very low, but our distance variable to be significant at at least the 0.01 level frequently. This lead us to believe that our regressions accounted for very little of the explanation for health outcomes across US counties, but distance still was an important contributor. Our regressions advanced to include smoking data, control by county, look at the closest 10 percent of county centroids to plants, and look at the top 50 percent of closest county centroids to plants. ###

```

## Advanced Regressions

```{r}
data <- final
```

```{r}
head(data)
colnames(data)
```

```{r}
library(dplyr)
library(magrittr)

data <- data %>% 
  mutate(distance_km = distance_to_nearest_plant / 1000)

# The file smoking_CLEAN provides more independent variables for control. The crucial variable provided by this data set is the crude percent of adults who currently smoke in 2020.

smoking <- read_csv(file = "https://raw.githubusercontent.com/UNCPublicPolicy/final-team-projects-darkorchid/main/DarkOrchid/DATA/Clean/smoking_CLEAN.csv")


#Making the fips code identical- changed geo_id to "fips" and used sprintf to add an extra "0" in front of codes that were 4 digits long isntead of 5. This made the fips codes compatible between the smoking data and the rest of the compiled data. 

smoking <- smoking %>% rename(fips = geo_id)

data <- data %>%
  mutate(fips = sprintf("%05d", fips))

# Join the two data sets using the FIPS code

merged_data <- data %>%
  left_join(smoking, by = c("fips", "county"))

merged_data <- merged_data %>%
  rename(state = state.x, state_abrv = state.y)

#file_path3 <- UNCPublicPolicy/final-team-projects-darkorchid/DATA/clean/health_aqi_coal_distance_smoking_CLEAN.csv

#write_csv(merged_data, file_path3)

```

```{r}
data_10 <- merged_data %>%
  arrange(distance_km) %>% # sort by distance_km
  slice(1:round(n() * 0.1)) # keep top 10% of observations

data_50 <- merged_data %>%
  arrange(distance_km) %>% # sort by distance_km
  slice(1:round(n() * 0.5)) # keep top 50% of observations
```

```{r}
#An experimental model including many different independent variables. Investigating realtionship for mortality_resp. 
exmodel <- lm(mortality_resp_2014 ~ median_aqi + max_aqi + good_days + moderate_days + unhealthy_for_sensitive_groups_days + unhealthy_days + hazardous_days +days_ozone + days_no2 + distance_km, data = merged_data)
summary(exmodel)
```

```{r}
# Experiments with mortality_resp models. 
exmodel2 <- lm(mortality_resp_2014 ~ median_aqi + unhealthy_days + hazardous_days + days_no2 + distance_km, data = merged_data)
summary(exmodel2)
```

```{r}
final_mod1 <-lm(mortality_resp_2014 ~ distance_km, data = merged_data)

summary(final_mod1)
```

```{r}
# A deeper look into distance_km's effect on mortality_resp. Only analyzing the top 10 percent lowest values for min_distance from caol fired power plant. 

final_mod10 <-lm(mortality_resp_2014 ~ distance_km, data = data_10)

summary(final_mod10)
```


```{r}
# A deeper look into distance_km's effect on mortality_resp. Only analyzing the top 50 percent lowest values for min_distance from caol fired power plant. 

final_mod50 <-lm(mortality_resp_2014 ~ distance_km, data = data_50)

summary(final_mod50)
```


```{r}
# Investigating the strength of current_smokers effect on moratlity_resp compared to distance_km. 

smoking_mod1 <-lm(mortality_resp_2014 ~ distance_km + current_smokers, data = merged_data)

summary(smoking_mod1)
```


```{r}
# Investigating the strength of current_smokers effect on moratlity_resp compared to distance_km controlling by state. 

smoking_mod2 <- lm(mortality_resp_2014 ~ distance_km + current_smokers + state, data = merged_data)

summary(smoking_mod2)
```


```{r}
# Investigating the strength of current_smokers effect on moratlity_resp compared to distance_km. Only analyzing the top 10 percent lowest values for min_distance from coal fired power plant. 
smoking_mod10 <- lm(mortality_resp_2014 ~ distance_km + current_smokers, data = data_10)

summary(smoking_mod10)
```


```{r}
# Investigating the strength of current_smokers effect on moratlity_resp compared to distance_km. Only analyzing the top 50 percent lowest values for min_distance from coal fired power plant. 
smoking_mod50 <- lm(mortality_resp_2014 ~ distance_km + current_smokers, data = data_50)

summary(smoking_mod50)
```

