---
title: "Dark Orchid EDA"
author: "Betsy Norwood"
date: "4/14/2023"
output: html_document
---

```{r}

setwd("/Users/betsynorwood/Documents/GitHub/final-team-projects-darkorchid/DarkOrchid/Output")
getwd()

library(tidyverse)
library(ggplot2)
library(dplyr)
library(readr)
library(naniar)
library(patchwork)

data <- read.csv("fulldata.csv")

```

fips: 
location: the location of the power plant listed as county, state  
county: the county of the power plant 

state: the state of the power plant 
mortality_resp: 
perc_mortality_change_resp:

mortality_ast: 
perc_mortality_change_ast: 
mortality_lung:

perc_morality_change_lung: 
year: 
days_with_aqi: 

good_days: 
moderate_days: 
unhealthy_for_sensative_groups_days: 

unhealthy_days: 
very_unhealthy_days: 
hazardous_days: 

max_aqi: 
x90th_percentile_aqi: 
median_aqi: 

days_co: 
days_no2:
days_ozone: 

days_pm2_5: 
days_pm10: 
lat: 

long: 
distance_to_nearest_plant: 
name_of_nearest_plant: 

```{r}

glimpse(data)
names(data)

data %>% summarise_all(funs(sum(is.na(.))))

vis_miss(data)

```

```{r, plots}

# max_aqi 
ggplot(data, aes(x=max_aqi)) +
  geom_density(alpha=0.5,show.legend = FALSE, color="darkorchid") +
  xlim(0, 500) + # scale for aqi is 0 (good) to 500 (hazardous)
  theme_minimal()

# morality_ast 
ggplot(data, aes(x=mortality_ast)) +
  geom_histogram(alpha=1)+
  theme_minimal()

# Patchwork of AQI Value Categories 
p1 <- ggplot(data, aes(x=good_days)) +
  geom_density(alpha=0.5,show.legend = FALSE, color="darkorchid") +
  theme_minimal()

p2 <- ggplot(data, aes(x=moderate_days)) +
  geom_density(alpha=0.5,show.legend = FALSE, color="darkorchid") +
  theme_minimal()

p3 <- ggplot(data, aes(x=unhealthy_for_sensitive_groups_days)) +
  geom_density(alpha=0.5,show.legend = FALSE, color="darkorchid") +
  theme_minimal()

p4 <- ggplot(data, aes(x=unhealthy_days)) +
  geom_density(alpha=0.5,show.legend = FALSE, color="darkorchid") +
  theme_minimal()

p5 <- ggplot(data, aes(x=very_unhealthy_days)) +
  geom_density(alpha=0.5,show.legend = FALSE, color="darkorchid") +
  theme_minimal()

p6 <- ggplot(data, aes(x=hazardous_days)) +
  geom_density(alpha=0.5,show.legend = FALSE, color="darkorchid") +
  theme_minimal()

p1 + p2 + p3+ p4 + p5 + p6
```